---
title: "Untitled"
author: "Ngodoo Zume"
date: "7/26/2020"
output: html_document
---
```{r, include=FALSE}
library(readxl)
state_data <- read_excel("state_data.xlsx") #already incudes everything below

pop <- data.frame(read.csv("covid_county_population_usafacts.csv"))

pop <- aggregate(pop$population, by=list(state=pop$State), FUN=sum, na.rm=TRUE)
names(pop)[2] <- "state_pop"


#combine with state data by state
state_data <- merge(state_data, pop, by = "state", all.x = TRUE)

#all NA in pct_chng column = 0 
state_data$pct_change[is.na(state_data$pct_change)] <- 0

#shorten percentage to 2 decimal places
state_data[,'pct_change'] <- format(round(state_data[,'pct_change'],2),nsmall=2)


voting_data <- read_excel("2018 Voting Data.xlsx")
names(voting_data)
state_data <- merge(state_data, voting_data, by = "state", all.x = TRUE)

write_xlsx(state_data, "state_data.xlsx")

#Sum all state level data  
state_cases <- aggregate(state_data$cases, by=list(state=state_data$state), FUN=sum)
names(state_cases)[2]<- "Cases"
state_cases<- state_cases[order(state_cases$state),]

state_population <- aggregate(state_data$state_pop, by=list(state=state_data$state), FUN=mean)
names(state_population)[2]<- "Population"
state_population <- state_population[order(state_population$state),]

state_chr <- aggregate(deaths~ state+party, state_data, sum)
state_chr <- state_chr[order(state_chr$state),]

#Merge datasets 
data_by_state <- cbind(state_cases, state_population, state_chr)
data_by_state <- data_by_state[,c(1,2,4,6,7)]
data_by_state

#Rename columns
names(data_by_state)[1]<- "State"
names(data_by_state)[4]<- "Party"
names(data_by_state)[5]<- "Deaths"

#Format Dataset
data_by_state$Party[data_by_state$Party== "rep"]<- "Republican"
data_by_state$Party[data_by_state$Party== "dem"]<- "Democratic"
data_by_state

data_by_state$Party <- factor(data_by_state$Party, levels= c("Republican", "Democratic"))
data_by_state$State <- factor(data_by_state$State)
data_by_state$Population <- as.integer(data_by_state$Population)
data_by_state

#Account for population
data_by_state <- transform(data_by_state, new= Population/1000)
names(data_by_state)[6]<- "P"
data_by_state$P <- as.integer(data_by_state$P)

data_by_state <- transform(data_by_state, new= Deaths / P)
names(data_by_state)[7]<- "D"
data_by_state$D <- as.integer(data_by_state$D)

data_by_state <- transform(data_by_state, new= Cases / P)
names(data_by_state)[8]<- "C"
data_by_state$C <- as.integer(data_by_state$C)

names(data_by_state)[6]<- "Pop.per.thou"
names(data_by_state)[7]<- "Deaths.per.thou"
names(data_by_state)[8]<- "Cases.per.thou"

write_xlsx(data_by_state, "data_by_state.xlsx")
```

```{r, include=FALSE}
sum <- state_data[c(-1, -2, -3, -10)]
by(sum, sum$party, summary)
```
```{r, include= FALSE}
#s <- y
#by(s, y$party, summary)

ag <- aggregate(. ~ party, y, function(x) c(sd = sd(x)))
ag
```

#1. Introduction


##1.1 Index
This report is organized as follows: 
1. Introduction
1.1 Index
2. Description of the Data
2.1 Source Data
2.2 Variables
2.3 Geographical Description of the Data
3. Analysis
3.1 Multiple Linear Regression Model
3.2 Two Sample T-Test
3.3 Linear Regression 
3.4 Anova 
4. Conclusion
5. Citations

#2. Description of the Data 
##2.1 Source Data

##2.2 Variables

##2.3 Geographical Description of the Data 

#3. Analysis

##3.2 Two Sample T-Test

```{r}
library(writexl)
write_xlsx(y, "y.xlsx")
write_xlsx(party_data, "party_data.xlsx")
```
```{r}
a <- log(y$avg_change)
hist(a)
```
```{r}
library(ggplot2)
ggplot(data=state_data, aes(x=reorder(state, -state_pop), y=state_pop, fill=party)) + ggtitle("state population") +
  xlab("State") + ylab("population") +
  geom_bar(stat="identity") + scale_fill_manual(values = c("blue", "red")) + theme(axis.text.x=element_text(angle=90, hjust=1), plot.title = element_text(hjust = 0.5,  size=14, face="bold"))
```

##3.3 Linear Regression 
```{r, include=FALSE}
#Remove outliers from cases and deaths per thousand 
data_by_state_clean<- outlierKD2(data_by_state, Cases.per.thou, TRUE)
data_by_state_clean<- outlierKD2(data_by_state, Deaths.per.thou, TRUE)
```
```{r, include=FALSE}
#Subset republican and dem data 
rep_data1 <- subset(data_by_state_clean, Party == "Republican")
dem_data1 <- subset(data_by_state_clean, Party == "Democratic")
```
```{r}
#linear model cases v deaths

library(ggplot2)
 
ggplot(data_by_state_clean, aes(x=Cases.per.thou, y=Deaths.per.thou, color=Party))+ geom_point()+labs(title="Scatterplot", subtitle = "Deaths vs Cases", x= "Cases (per 1,000)", y= "Deaths (per 1,000)") +geom_smooth(method = "lm" )+ scale_color_manual(values=c("red", "blue"))+ xlim(c(0,850))+ ylim(c(0,50))

```
Next, we wanted to look at the relationship between COVID related deaths and the number of cases. More specifically, do deaths trend with cases and what does that trend look like for each party? In this section we are factoring in population by the thousands and have removed outliers to decrease variability and increase statistical power. To analyze this relationship, we created a scatterplot formatted with a line of best fit. The line of best fit follows the equation ŷ = bX + a, where b is the slope of the line and a is the intercept, which allows us to estimate the dependent variable (deaths), given the independent variable (cases). We would assume, as a state’s number of cases  increases, so should its number of deaths, providing a positive trend, which is observed in our model. This graph shows an increased number of democratic states with higher case and death occurrences. We also see that the rate of deaths per case for both parties look similar. 

```{r}
#Correlation
cor.test(rep_data1$Cases.per.thou, rep_data1$Deaths.per.thou, method = "pearson", conf.level = 0.95)
```
```{r}
cor.test(dem_data1$Cases.per.thou, dem_data1$Deaths.per.thou, method = "pearson", conf.level = 0.95)
```
To analyze the relationship between cases and deaths among both parties we completed a Pearson's correlation test, which measures the strength of a linear association between two variables, at a 95% confidence interval. When we look at the level of linear dependence between our two variables, both republicans (r=.726) and democrats (r=.966) have a strong correlation. Democrats correlation between deaths and cases is higher. 

Next, we completed a linear regression model to see if we can predict the value of an outcome (deaths) based on our predictor variable (cases). Our null hypothesis is that there is no relationship between the independent variable (deaths) and our dependent variable (cases). The alternative hypothesis is that the coefficients are not equal to zero.

```{r}
rep_lm <- glm(formula= Cases.per.thou~Deaths.per.thou, data=rep_data1)
summary(rep_lm)
```
```{r}
dem_lm <- glm(formula= Cases.per.thou~Deaths.per.thou , data=dem_data1)
summary(dem_lm)
```
Through these tests we see a statistical significance (p<.05) between these two linear models for republicans and democrats, allowing us to reject the null hypothesis that there is no relationship between cases and deaths for each party. We also see that republican (b=11.88 ) and democratic (b= 20.56) states have differing slopes, suggesting that the rate of deaths and cases for parties vary. 
```{r}
confint(rep_lm)
```
```{r}
confint(dem_lm)
```
To test if this difference was significant, we ran a confidence interval for each set of data (Figure 12-13). As the intervals lie just outside of each other, we can conclude with 95% confidence that the deaths and cases for each party are different. 

##3.4 ANOVA

#4. Conclusion

#5. Citations 
